### 2010-09-21 doc/guide/options.tex
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/trunk/gringo@3396 3e02f596-3131-0410-8352-e04eeaec37cc

In this section, we briefly describe the meanings of
command line options supported by
\gringo\ (Section~\ref{subsec:opt:gringo}),
\clingo\ (Section~\ref{subsec:opt:clingo}),
\iclingo\ (Section~\ref{subsec:opt:iclingo}), and
\clasp\ (Section~\ref{subsec:opt:clasp}).
Each of these tools displays its available options
when invoked with flag \code{--help} or \code{-h},
respectively.%
\footnote{%
  Note that our description of command line options
  is based on Version~2.0.2 of \gringo, \clingo, and \iclingo\
  as well as Version~1.1.2 of \clasp.
  While it is rather unlikely that command line options will
  disappear in future versions,
  additional ones might be introduced.
  We will try to keep this document up-to-date,
  but checking the help information shipped 
  with a new version is always a good idea.}
The approach of distinguishing long options, starting with ``\code{--},''
and short ones of the form ``\code{-\textit{l}},''
where \code{\textit{l}} is a letter,
follows the GNU Coding Standards~\cite{GNUcoding}.
For obvious reasons,
short forms are made available only for the most common (long) options.
Some options, also called flags, do not take any argument,
while others require arguments.
An argument \code{\textit{arg}} is provided to a (long) option \code{\textit{opt}}
by writing 
``\code{--\textit{opt}=\textit{arg}}'' or
``\code{--\textit{opt} \textit{arg}},''
while only
``\code{-\textit{l} \textit{arg}}''
is accepted for a short option \code{\textit{l}}.
For each command line option,
we below indicate whether it requires an argument,
and if so, we also describe its meaning.

@@

An abstract invocation of \gringo\ looks as follows:
%
\begin{lstlisting}[numbers=none]
gringo [ options | filenames ]
\end{lstlisting}
%
Note that options and filenames do not need to be passed to \gringo\
in any particular order.
If neither a filename nor an option that makes \gringo\ exit (see below)
is provided, \gringo\ reads from the standard input.
In the following, we list and describe the options accepted by \gringo\
along with their particular arguments (if required):
%
\begin{description}
\item[\code{--help,-h}]~\\
Print help information and exit.
\item[\code{--version,-v}]~\\
Print version information and exit.
%\item[\code{--syntax}]~\\
%Print syntax information (about the input language) and exit.
%\item[\code{--stats}]~\\
%Print (extended) statistic information before termination.
\item[\code{--verbose[=n],-V=[n]}]~\\
Print additional (progress) information during computation.
Verbosity level one and two is reserved for clasp only.
Level three prints internal representations of rules during grounding.
(This may be used to identify either semantic errors in an input program or
 performance bottlenecks.)
\item[\code{--const,-c \textit{c}=\textit{t}}]~\\
Replace occurrences (in the input program) of a constant~\code{\textit{c}}
with a term~\code{\textit{t}}.
\item[\code{--text,-t}]~\\
Output ground program in (human-readable) text format.
\item[\code{--reify}]~\\
Output ground program in form of facts.
(These facts can then be used to, i.e., write a meta-interpreter.)
\item[\code{--lparse,-l}]~\\
Output ground program in \lparse's numerical format~\cite{lparseManual}.
\item[\code{--ground,-g}]~\\
Enable lightweight mode for processing a ground input program.
(This option is recommended to 
 omit unnecessary overhead if the input program is already ground,
 but it leads to a syntax error (cf.\ Section~\ref{subsec:error}) otherwise.)
\item[\code{--ifixed \textit{n}}]~\\
Use \code{\textit{n}} as fix step number if the input program
contains \const{\#cumulative} or \const{\#volatile} statements.
(This option permits the handling of programs written for \iclingo\
 in a traditional single pass computation.)
\item[\code{--ibase}]~\\
Process only the static part (that can be initiated by a \const{\#base} statement)
of an input program.
(This option may be used to investigate the basic setting of a problem
 including some mutable bound.)
\item[\code{--dep-graph=filename}]~\\
This option can be used to get the dependency graph in from
of a dot\footnote{\url{http://www.graphviz.org/}} file of the program.
\item[\code{--shift}]~\\
Removes disjunction by shifting (see \cite{gellif91a}).
\end{description}
%
The default command line when invoking \gringo\ is as follows:
%
\begin{lstlisting}[numbers=none]
gringo --lparse
\end{lstlisting}
That is, \gringo\ usually outputs a ground program in \lparse's numerical format,
dealt with by various ASP solvers \cite{gekanesc07b,siniso02a,linzha04a}.

@@

We below group general options of \clasp,
used to configure its global behavior.
%
\begin{description}
\item[\code{--help,-h}]~\\
Print help information and exit.
\item[\code{--version,-v}]~\\
Print version information and exit.
\item[\code{--stats}]~\\
Print (extended) statistic information before termination.
\item[\code{--file,-f \textit{filename}}]~\\
Read from file \code{\textit{filename}} (rather than from the standard input).
\item[\code{--number,-n \textit{n}}]~\\
Compute at most~\code{\textit{n}} answer sets,
$\code{\textit{n}}=\code{0}$ standing for compute all answer sets.
\item[\code{--quiet,-q}]~\\
Do not print computed answer sets.
(This is useful for benchmarking.)
\item[\code{--seed \textit{n}}]~\\
Use seed~\code{\textit{n}} (rather than~\code{1}) for random number generator.
\item[\code{--brave}]~\\
Compute the brave consequences (union of all answer sets) of a logic program.
\item[\code{--cautious}]~\\
Compute the cautious consequences (intersection of all answer sets) of a logic program.
\item[\code{--opt-all}]~\\
Compute all optimal answer sets (cf.\ Section~\ref{subsec:gringo:optimize}).
(This is implemented by enumerating~\cite{gekanesc07c} answer sets that are not worse
 than the best one found so far.)
\item[\code{--opt-restart}]~\\
Restart the search from scratch (rather than doing enumeration~\cite{gekanesc07c})
after finding a provisionally optimal answer set.
(This may speed up finding the optimum.)
\item[\code{--opt-value \textit{n1}[,\textit{n2},\textit{n3}...]}]~\\
Initialize objective function(s) to minimize with \code{\textit{n1}[,\textit{n2},\textit{n3}...]}.
Depending on the reasoning mode,
only equitable or strictly better answer sets are computed.
\item[\code{--supp-models}]~\\
Compute supported models~\cite{apblwa88a} (rather than answer sets).
\item[\code{--dimacs}]~\\
Compute models of a propositional formula in DIMACS/CNF format~\cite{dimacs}.
\item[\code{--trans-ext all|choice|weight|no}]~\\
Compile extended rules~\cite{siniso02a} into normal rules of form (\ref{eq:normal:rule}).
Arguments \code{choice} and \code{weight}
state that all ``choice rules'' or all ``weight rules,'' respectively,
are to be compiled into normal rules,
while \code{all} means that both and \code{no} that none of them
are subject to compilation.
\item[\code{--eq \textit{n}}]~\\
Run equivalence reasoning~\cite{gekanesc08a} for \code{\textit{n}} iterations,
$\code{\textit{n}}=\code{-1}$ and $\code{\textit{n}}=\code{0}$ standing for
run to fixpoint or do not run equivalence reasoning, respectively.
\item[\code{--sat-prepro yes|no|\textit{n1}[,\textit{n2},\textit{n3}]}]~\\
Run \emph{SatElite}-like preprocessing~\cite{eenbie05a}
for at most~\code{\textit{n1}} iterations 
($\code{\textit{n1}}=\code{-1}$ standing for run to fixpoint),
using cutoff~\code{\textit{n2}} for variable elimination
($\code{\textit{n2}}=\code{-1}$ standing for no cutoff), and
for no longer than~\code{\textit{n3}} seconds
($\code{\textit{n3}}=\code{-1}$ standing for no time limit).
Arguments \code{yes} and \code{no} mean
$\code{\textit{n1}}=\code{\textit{n2}}=\code{\textit{n3}}=\code{-1}$
(that is, run to fixpoint)
or that \emph{SatElite}-like preprocessing is not to be run at all, respectively.
\item[\code{--rand-watches yes|no}]~\\
Initially choose watched literals randomly (with argument \code{yes})
or systematically (with argument \code{no}).
\end{description}
%
Having introduced the general options of \clasp,
let us note that the options below \code{--dimacs} in the above list
are quite low-level and more or less an issue of fine-tuning.
More important is the fact that virtually all optimization functionalities
are only provided by \clasp\ if the maximum number of answer sets to be
computed is set to~\code{0} (standing for all answer sets),
as it is likely to stop search too early otherwise.
The same applies to computing either brave or cautious consequences
(via one of the flags \code{--brave} and \code{--cautious}).

@@

The options listed below can be used to configure the main search
strategies of \clasp.
%
\begin{description}
\item[\code{--lookback yes|no}]~\\
Enable or disable lookback techniques (cf.\ Section~\ref{subsec:clasp:back}).
(This option is included mainly for comparison purposes, and generally,
 it is recommended to keep the default argument value \code{yes}.)
\item[\code{--lookahead atom|body|hybrid|auto|no}]~\\
Apply failed-literal detection~\cite{freeman95}
to atoms (with argument \code{atom}),
to rule bodies (with argument \code{body}),
to atoms and rule bodies like in \nomorepp~\cite{angelinesc05c} (with argument \code{hybrid}),
or let \clasp\ pick one of the previous three possibilities
based on program properties (with argument \code{auto}).
Failed-literal detection is switched off via argument~\code{no}.
\item[\code{--initial-lookahead}]~\\
Apply failed-literal detection (to atoms) in preprocessing,
but not (necessarily) during search.
\item[\code{--heuristic Berkmin|Vmtf|Vsids|Unit|None}]~\\
Use \emph{BerkMin}-like decision heuristic~\cite{golnov02a}
(with argument~\code{Berkmin}),
\emph{Siege}-like decision heuristic~\cite{ryan04a}
(with argument~\code{Vmtf}),
\emph{Chaff}-like decision heuristic~\cite{momazhzhma01a}
(with argument~\code{Vsids}),
\emph{Smodels}-like decision heuristic~\cite{siniso02a}
(with argument~\code{Unit}),
or (arbitrary) static variable ordering
(with argument~\code{None}).
\item[\code{--rand-freq \textit{p}}]~\\
Perform random (rather than heuristic) decisions
with probability $0\leq\code{\textit{p}}\leq 1$.
\item[\code{--rand-prob yes|no|\textit{n1},\textit{n2}}]~\\
Run \emph{Satzoo}-like random probing~\cite{een03a},
initially performing~\code{\textit{n1}} passes of up to~\code{\textit{n2}} conflicts
making random decisions.
Arguments \code{yes} and \code{no} mean
$\code{\textit{n1}}=\code{50},\linebreak[1]\code{\textit{n2}}=\code{20}$
or that random probing is not to be run at all, respectively.
\end{description}
%
Let us note that switching the above options can have dramatic effects
(both positively and negatively) on the search performance of \clasp.
Sticking to the default argument values \code{yes} for \code{--lookback}
and \code{no} for \code{--lookahead} is generally recommended.
If nonetheless performance bottlenecks are observed,
it is worthwhile to give \code{Vmtf} and \code{Vsids} for \code{--heuristic}
a try, in particular,
when the program under consideration is huge 
but scarcely yields conflicts during search.

>>>>

In this section, we briefly describe the meanings of
command line options supported by
\gringo\ (Section~\ref{subsec:opt:gringo}),
\clingo\ (Section~\ref{subsec:opt:clingo}),
\iclingo\ (Section~\ref{subsec:opt:iclingo}), and
\clasp\ (Section~\ref{subsec:opt:clasp}).
Each of these tools displays its available options
when invoked with flag \code{--help} or \code{-h},
respectively.%
\footnote{%
  Note that our description of command line options
  is based on Version~3.0.x of \gringo, \clingo, and \iclingo\
  as well as Version~1.3.x of \clasp.
  While it is rather unlikely that command line options will
  disappear in future versions,
  additional ones might be introduced.
  We will try to keep this document up-to-date,
  but checking the help information shipped 
  with a new version is always a good idea.}
The approach of distinguishing long options, starting with ``\code{--},''
and short ones of the form ``\code{-\textit{l}},''
where \code{\textit{l}} is a letter,
follows the GNU Coding Standards~\cite{GNUcoding}.
For obvious reasons,
short forms are made available only for the most common (long) options.
Some options, also called flags, do not take any argument,
while others require arguments.
An argument \code{\textit{arg}} is provided to a (long) option \code{\textit{opt}}
by writing 
``\code{--\textit{opt}=\textit{arg}}'' or
``\code{--\textit{opt} \textit{arg}},''
while only
``\code{-\textit{l} \textit{arg}}''
is accepted for a short option \code{\textit{l}}.
For each command line option,
we below indicate whether it requires an argument,
and if so, we also describe its meaning.

@@

An abstract invocation of \gringo\ looks as follows:
%
\begin{lstlisting}[numbers=none]
gringo [ options | filenames ]
\end{lstlisting}
%
Note that options and filenames do not need to be passed to \gringo\
in any particular order.
If neither a filename nor an option that makes \gringo\ exit (see below)
is provided, \gringo\ reads from the standard input.
In the following, we list and describe the options accepted by \gringo\
along with their particular arguments (if required):
%
\begin{description}
\item[\code{--help,-h}]~\\
Print help information and exit.
\item[\code{--version,-v}]~\\
Print version information and exit.
%\item[\code{--syntax}]~\\
%Print syntax information (about the input language) and exit.
%\item[\code{--stats}]~\\
%Print (extended) statistic information before termination.
\item[\code{--verbose[=n],-V=[n]}]~\\
Print additional (progress) information during computation.
Verbosity level one and two is reserved for clasp only.
Level three prints internal representations of rules during grounding.
(This may be used to identify either semantic errors in an input program or
 performance bottlenecks.)
\item[\code{--const,-c \textit{c}=\textit{t}}]~\\
Replace occurrences (in the input program) of a constant~\code{\textit{c}}
with a term~\code{\textit{t}}.
\item[\code{--text,-t}]~\\
Output ground program in (human-readable) text format.
\item[\code{--reify}]~\\
Output ground program in form of facts.
(These facts can then be used to, i.e., write a meta-interpreter.)
\comment{add a section about a meta-interpreter}
\item[\code{--lparse,-l}]~\\
Output ground program in \lparse's numerical format~\cite{lparseManual}.
\item[\code{--ground,-g}]~\\
Enable lightweight mode for processing a ground input program.
(This option is recommended to 
 omit unnecessary overhead if the input program is already ground,
 but it leads to a syntax error (cf.\ Section~\ref{subsec:error}) otherwise.)
\item[\code{--ifixed \textit{n}}]~\\
Use \code{\textit{n}} as fix step number if the input program
contains \const{\#cumulative} or \const{\#volatile} statements.
(This option permits the handling of programs written for \iclingo\
 in a traditional single pass computation.)
\item[\code{--ibase}]~\\
Process only the static part (that can be initiated by a \const{\#base} statement)
of an input program.
(This option may be used to investigate the basic setting of a problem
 including some mutable bound.)
\item[\code{--dep-graph=filename}]~\\
This option can be used to get the dependency graph in from
of a dot\footnote{\url{http://www.graphviz.org/}} file of the program.
\item[\code{--shift}]~\\
Removes disjunction by shifting (see \cite{gellif91a}).
\end{description}
%
The default command line when invoking \gringo\ is as follows:
%
\begin{lstlisting}[numbers=none]
gringo --lparse
\end{lstlisting}
That is, \gringo\ usually outputs a ground program in \lparse's numerical format,
dealt with by various ASP solvers \cite{gekanesc07b,siniso02a,linzha04a}.

@@

We below group general options of \clasp,
used to configure its global behavior.
%
\begin{description}
\item[\code{--help,-h}]~\\
Print help information and exit.
\item[\code{--version,-v}]~\\
Print version information and exit.
\item[\code{--stats}]~\\
Print (extended) statistic information before termination.
\item[\code{--file,-f \textit{filename}}]~\\
Read from file \code{\textit{filename}} (rather than from the standard input).
\item[\code{--number,-n \textit{n}}]~\\
Compute at most~\code{\textit{n}} answer sets,
$\code{\textit{n}}=\code{0}$ standing for compute all answer sets.
\item[\code{--quiet,-q}]~\\
Do not print computed answer sets.
(This is useful for benchmarking.)
\item[\code{--seed \textit{n}}]~\\
Use seed~\code{\textit{n}} (rather than~\code{1}) for random number generator.
\item[\code{--brave}]~\\
Compute the brave consequences (union of all answer sets) of a logic program.
\item[\code{--cautious}]~\\
Compute the cautious consequences (intersection of all answer sets) of a logic program.
\item[\code{--opt-all}]~\\
Compute all optimal answer sets (cf.\ Section~\ref{subsec:gringo:optimize}).
(This is implemented by enumerating~\cite{gekanesc07c} answer sets that are not worse
 than the best one found so far.)
\item[\code{--opt-restart}]~\\
Restart the search from scratch (rather than doing enumeration~\cite{gekanesc07c})
after finding a provisionally optimal answer set.
(This may speed up finding the optimum.)
\item[\code{--opt-value \textit{n1}[,\textit{n2},\textit{n3}...]}]~\\
Initialize objective function(s) to minimize with \code{\textit{n1}[,\textit{n2},\textit{n3}...]}.
Depending on the reasoning mode,
only equitable or strictly better answer sets are computed.
\item[\code{--supp-models}]~\\
Compute supported models~\cite{apblwa88a} (rather than answer sets).
\item[\code{--trans-ext all|choice|weight|no}]~\\
Compile extended rules~\cite{siniso02a} into normal rules of form (\ref{eq:normal:rule}).
Arguments \code{choice} and \code{weight}
state that all ``choice rules'' or all ``weight rules,'' respectively,
are to be compiled into normal rules,
while \code{all} means that both and \code{no} that none of them
are subject to compilation.
\item[\code{--eq \textit{n}}]~\\
Run equivalence reasoning~\cite{gekanesc08a} for \code{\textit{n}} iterations,
$\code{\textit{n}}=\code{-1}$ and $\code{\textit{n}}=\code{0}$ standing for
run to fixpoint or do not run equivalence reasoning, respectively.
\item[\code{--sat-prepro yes|no|\textit{n1}[,\textit{n2},\textit{n3}]}]~\\
Run \emph{SatElite}-like preprocessing~\cite{eenbie05a}
for at most~\code{\textit{n1}} iterations 
($\code{\textit{n1}}=\code{-1}$ standing for run to fixpoint),
using cutoff~\code{\textit{n2}} for variable elimination
($\code{\textit{n2}}=\code{-1}$ standing for no cutoff), and
for no longer than~\code{\textit{n3}} seconds
($\code{\textit{n3}}=\code{-1}$ standing for no time limit).
Arguments \code{yes} and \code{no} mean
$\code{\textit{n1}}=\code{\textit{n2}}=\code{\textit{n3}}=\code{-1}$
(that is, run to fixpoint)
or that \emph{SatElite}-like preprocessing is not to be run at all, respectively.
\item[\code{--rand-watches yes|no}]~\\
Initially choose watched literals randomly (with argument \code{yes})
or systematically (with argument \code{no}).
\end{description}
%
Having introduced the general options of \clasp,
let us note that the options below \code{--supp-models} in the above list
are quite low-level and more or less an issue of fine-tuning.
More important is the fact that virtually all optimization functionalities
are only provided by \clasp\ if the maximum number of answer sets to be
computed is set to~\code{0} (standing for all answer sets),
as it is likely to stop search too early otherwise.
The same applies to computing either brave or cautious consequences
(via one of the flags \code{--brave} and \code{--cautious}).

@@

The options listed below can be used to configure the main search
strategies of \clasp.
%
\begin{description}
\item[\code{--lookback yes|no}]~\\
Enable or disable lookback techniques (cf.\ Section~\ref{subsec:clasp:back}).
(This option is included mainly for comparison purposes, and generally,
 it is recommended to keep the default argument value \code{yes}.)
\item[\code{--lookahead atom|body|hybrid|auto|no}]~\\
Apply failed-literal detection~\cite{freeman95}
to atoms (with argument \code{atom}),
to rule bodies (with argument \code{body}),
to atoms and rule bodies like in \nomorepp~\cite{angelinesc05c} (with argument \code{hybrid}),
or let \clasp\ pick one of the previous three possibilities
based on program properties (with argument \code{auto}).
Failed-literal detection is switched off via argument~\code{no}.
\item[\code{--initial-lookahead}]~\\
Apply failed-literal detection (to atoms) in preprocessing,
but not (necessarily) during search.
\item[\code{--heuristic Berkmin|Vmtf|Vsids|Unit|None}]~\\
Use \emph{BerkMin}-like decision heuristic~\cite{golnov02a}
(with argument~\code{Berkmin}),
\emph{Siege}-like decision heuristic~\cite{ryan04a}
(with argument~\code{Vmtf}),
\emph{Chaff}-like decision heuristic~\cite{momazhzhma01a}
(with argument~\code{Vsids}),
\emph{Smodels}-like decision heuristic~\cite{siniso02a}
(with argument~\code{Unit}),
or (arbitrary) static variable ordering
(with argument~\code{None}).
\item[\code{--rand-freq \textit{p}}]~\\
Perform random (rather than heuristic) decisions
with probability $0\leq\code{\textit{p}}\leq 1$.
\item[\code{--rand-prob yes|no|\textit{n1},\textit{n2}}]~\\
Run \emph{Satzoo}-like random probing~\cite{een03a},
initially performing~\code{\textit{n1}} passes of up to~\code{\textit{n2}} conflicts
making random decisions.
Arguments \code{yes} and \code{no} mean
$\code{\textit{n1}}=\code{50},\linebreak[1]\code{\textit{n2}}=\code{20}$
or that random probing is not to be run at all, respectively.
\end{description}
%
Let us note that switching the above options can have dramatic effects
(both positively and negatively) on the search performance of \clasp.
Sticking to the default argument values \code{yes} for \code{--lookback}
and \code{no} for \code{--lookahead} is generally recommended.
If nonetheless performance bottlenecks are observed,
it is worthwhile to give \code{Vmtf} and \code{Vsids} for \code{--heuristic}
a try, in particular,
when the program under consideration is huge 
but scarcely yields conflicts during search.
For a brief overview about fine-tuning see \cite{gekasc09b}.
